---
title: "Penguins size comparisons"
author: "Michelle Shteyn"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```


# Part 1. T-tests

Mass comparison: male & female chinstrap penguins

```{r}
chinstraps<-penguins %>% 
  filter(species == "Chinstrap")

# Histograms
ggplot(data = chinstraps, aes(x=body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

#here you can see the mean body mass for males is closer to 4000, while for females it's closer to 3500


# QQ plot
ggplot(data = chinstraps, aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~sex)

#these are as close to normally distributed as you can hope real-world data will actually get

## Get descriptive statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )

#median is similar to mean, which reflects that symmetric distribution we saw in the histograms

```


Get vectors of body_mass_g for female and male chinstrap penguins separately.

```{r}
chinstraps_f<-chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m<-chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```


### T-tests, a couple of different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstrap_ttest<- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f,chinstraps_m)
```

The p value is the probability that I could have gotten sample means that are at least as different as ours are by random chance if they were drawn from populations with the same mean.


Are male chinstrap penguins significantly larger (body mass) than female chinstrap penguins?

```{r}
t.test(chinstraps_m,chinstraps_f, alternative = "greater")
```
Yes, I will reject the null hypothesis that the means difference is not greater than 0 and retain the alternative hypothesis that the means difference in body mass is greater than zero.


## One-way ANOVA

### 1. Table of counts by species and sex

```{r}
penguin_counts<-penguins %>% 
  count(species, sex)

penguin_counts
```


Create a subset of only female penguins
```{r}
penguin_f<-penguins %>% 
  filter(sex == "female")

ggplot(data = penguin_f, aes(x=flipper_length_mm))+
  geom_histogram(aes(fill=species))+
  facet_wrap(~species, scales = "free")


ggplot(data = penguin_f, aes(sample=flipper_length_mm))+
  geom_qq(aes(color=species))+
  facet_wrap(~species, scales = "free")
```

Both the histograms and qqplots tell us that these are very close to normally distributed groups.


### Summary statistics

Starting from penguins_f, create a summary table that groups by species and returns the following:

- mean of flipper length  
- sd ""  
- sample size  
- variance 'var'  
- standard error = sd/sqrt(n)

```{r}
penguin_f_summary<-penguin_f %>% 
  group_by(species) %>% 
  summarize(
    mean_flipper_length = mean(flipper_length_mm),
    sd_flipper_length = sd(flipper_length_mm),
    sample_size = n(),
    se_flipper_length = sd(flipper_length_mm)/sqrt(n()),
    var_flipper_length = var(flipper_length_mm)
    
  )

leveneTest(flipper_length_mm~species, data = penguin_f)
```


## Beeswarm plot with mean +- 1sd

```{r}
ggplot() +
  geom_beeswarm(data = penguin_f, 
                aes(x=species, y=flipper_length_mm),
                color = "gray70",
                size = 1,
                alpha = 0.5) + 
  geom_point(data = penguin_f_summary, aes(x=species, y=mean_flipper_length),
             size=4) +
  geom_errorbar(data = penguin_f_summary,
                aes(x = species,
                    ymax=mean_flipper_length + sd_flipper_length,
                    ymin = mean_flipper_length - sd_flipper_length),
                width=0.1)+
  theme_light()

```

## One-way ANOVA

```{r}
penguin_aov<- aov(flipper_length_mm ~ species, data = penguin_f)
summary(penguin_aov)

TukeyHSD(penguin_aov) 
```

The F-statistic is the ratio of the Bss / Wss. This is a very large F-statistic, and this is reflected in the p-value.

It is extremely unlikely that we could have taken samples from populations with the same mean and ended up with means that are at least this different.


If you have very unbalanced groups, you will want to do a type II ANOVA.

Type III Anova with the care package:
```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data=penguin_f)
penguin_aov_t3<-Anova(penguin_lm, type = "III")
penguin_aov_t3
```

If you have very different sample sizes, you will want to use a Type III anova. If it is very balanced, you can use a regular ANOVA.

There is definitely enough evidence to reject the null hypothesis that our samples were drawn from populations with the same mean.






































































































