---
title: "Penguins size comparisons"
author: "Michelle Shteyn"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```


# Part 1. T-tests

Mass comparison: male & female chinstrap penguins

```{r}
chinstraps<-penguins %>% 
  filter(species == "Chinstrap")

# Histograms
ggplot(data = chinstraps, aes(x=body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

#here you can see the mean body mass for males is closer to 4000, while for females it's closer to 3500


# QQ plot
ggplot(data = chinstraps, aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~sex)

#these are as close to normally distributed as you can hope real-world data will actually get

## Get descriptive statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )

#median is similar to mean, which reflects that symmetric distribution we saw in the histograms

```


Get vectors of body_mass_g for female and male chinstrap penguins separately.

```{r}
chinstraps_f<-chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m<-chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```


### T-tests, a couple of different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstrap_ttest<- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f,chinstraps_m)
```

The p value is the probability that I could have gotten sample means that are at least as different as ours are by random chance if they were drawn from populations with the same mean.


Are male chinstrap penguins significantly larger (body mass) than female chinstrap penguins?

```{r}
t.test(chinstraps_m,chinstraps_f, alternative = "greater")
```
Yes, I will reject the null hypothesis that the means difference is not greater than 0 and retain the alternative hypothesis that the means difference in body mass is greater than zero.





































